\documentclass{book}




\usepackage{fullpage}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{framed}
\usepackage{array}
\usepackage{algorithmic}
\usepackage{xcolor}

\definecolor{dark_red}{rgb}{0.5,0.0,0.0}
\definecolor{dark_green}{rgb}{0.0,0.5,0.0}
\definecolor{dark_blue}{rgb}{0.0,0.0,0.5}
\definecolor{red}{rgb}{1.0,0.0,0.0}
\definecolor{blue}{rgb}{0.0,0.0,1.0}

\newcommand{\dr}[1]{\textcolor{dark_red}{#1}}
\newcommand{\dg}[1]{\textcolor{dark_green}{#1}}
\newcommand{\db}[1]{\textcolor{dark_blue}{#1}}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\newcommand{\blue}[1]{\textcolor{blue}{#1}}

\newtheorem{lma}{Lemma}
\newtheorem{thm}{Theorem}
\newtheorem{cor}{Corollary}

\title{An Intuitive and Visual Approach to Vector Calculus and Electromagnetism}
\author{Shawn Eastwood}
\date{}

\begin{document}

\frontmatter

\maketitle


\chapter*{Introduction}

\addcontentsline{toc}{chapter}{Introduction}

Many discussions and introductions to vector calculus utilize examples and proofs that are exercises in symbolic computation. It is of this author's opinion that a symbolic approach to vector calculus does not convey core concepts in a sufficiently clear manner. Abstract proofs, abstract formulas, and examples for which the emphasis is on the algebra and symbolic computation do not convey the core essence of vector calculus.  

The primary purpose of this document is to provide an introduction to vector calculus and electromagnetism {\bf without using the language of Calculus}. The material presented below will use easy to visualize concepts and keep equations and numbers to a minimum. Any equation given will be accompanied by a high level description that will make such equations intuitive. 

The primary application that will be presented will be electromagnetism and Maxwell's equations. Most approaches to electromagnetism are piece-wise. For example, basic circuit analysis is primarily concerned with voltages, currents, and resistances. Eventually inductance and mutual inductance is introduced. What is rarely mentioned in these discussions is that the electric field is not conservative, and therefore the notion of voltage itself is poorly defined. While the magnetic vector potential can be used to ``complete" the electric field and create a field that is conservative, the vector potential itself is not uniquely defined and is subject to multiple ``gauges". Within inductors in particular, the voltage difference across the inductor is in direct opposition to the electric field that actually exists. The voltage difference is instead the electric field that must exist to ``push back" the electric field generated by the inductor itself.

%Perfect current itself does not actually exist: current is the movement of point charges, mostly electrons. With the a more accurate picture of current, Maxwell's equations will be modified to a more accurate picture where the movement of field lines, not electric

The aim is to enable a visualization of electromagnetic phenomena that takes full account of Maxwell's equations. To this end, conventional depictions of vector fields as a forest of arrows will be replaced in favor of depictions that use bundles of oriented paths and layers of oriented surfaces. Using terminology from fluid mechanics, the description will be {\bf Lagrangian} in nature as opposed to more conventional {\bf Eulerian} descriptions. With Eulerian descriptions, the fields are tracked at {\bf fixed points}. With the Lagrangian description, the field lines will be tracked as they sweep through space. 

\section*{What will be assumed}

This document will assume that the reader is familiar with basic algebra. 

While a basic knowledge of Calculus is useful, an in-depth knowledge of Calculus is not required. 

Since the conventional depiction of vectors as arrows is being replaced, vector algebra is also not required as a prerequisite topic. When vectors are introduced, they will be introduced as lists of numbers.

\section*{Chapter structure}

Chapter 1 introduces the fundamental structures of points, paths, surfaces, and volumes. Chapters 2 to 4 introduce the basic functions involving points, paths, surfaces, and volumes. Chapters 5 and 6 establish algebraic properties involving the introduced structures and operations. 

Chapter 7 introduces the system by which points, paths, surfaces, and volumes can be quantified directly with numbers and functions. It is in this chapter that a prior knowledge of differential calculus is useful. 

Chapter 8 describes how the dimensionality of structures can be ``inverted", with points becoming volumes and vice versa, and with paths becoming surfaces and vice versa. It is at this point that electromagnetism is introduced assuming no changes with respect to time (i.e. statics). Many important details related to electrostatics and magnetostatics will be proven. 

Chapter 9 finally introduces time. Additional functions are introduced. The computations from chapter 7 and the concept of duality from chapter 8 are extended to incorporate time.

Chapter 10 finally fully introduces electromagnetism with the incorporation of time.  

Chapter 11 connects electromagnetism with force, momentum, and energy. 


\tableofcontents




\mainmatter



\input{chapter_multi-structures.tex}



\input{chapter_unions.tex}



\input{chapter_intersections.tex}



\input{chapter_boundaries.tex}



\input{chapter_closed_loops_and_surfaces.tex}



\input{chapter_boundaries_and_intersections.tex}



\input{chapter_quantifying_multi-structures.tex}



\input{chapter_duality.tex}



\input{chapter_finite_element_modeling.tex}



\chapter{Introducing time}





\end{document}













